---
title: "Student Alcohol Consumption"
author: "Sachie Tran"
date: "2/14/2022"
output: html_document
---

## Introduction
<font size="4">

|       The data "student alcohol consumption" is collected from secondary school students who take both/either math and/or Portuguese language courses in Portugal. This includes over 1000 samples from various aspects like social, family, personal lifestyle and study status. The main purpose of this case study is to find how alcohol consumption relates to students, and how to control it for their academic goals, hoping to help young people and their family members achieve students' academic goals.

The original data is available from [Kaggle](https://www.kaggle.com/uciml/student-alcohol-consumption?select=student-mat.csv).

---

## Data Explore

|    <font size="5">- Summary of data
```{r load_packages, message=FALSE, warning=FALSE}
#install.packages("tidyverse")
#library(tidyverse)
#install.packages("dplyr")
library(dplyr)
#install.packages("reshape2")
library(reshape2)
```

```{r load_data}
students <- read.csv("all_students.csv")
head(students)
```

<font size="4">


<br/>


First, I would go through simple aggregations to get to know the data.

|         <font size="5">- Gender ratio
```{r gender_table}
Sex_tbl <- table(students$sex)  #F=Female, M=Male
Sex_tbl <- prop.table(Sex_tbl)*100
Sex_tbl <- round(Sex_tbl,digits=2)
Sex_tbl
```

|         <font size="5">- Age and gender ratio
```{r gender_sex_table}
AgeSex_tbl <- table(students$age, students$sex)
AgeSex_tbl <- prop.table(AgeSex_tbl)*100
AgeSex_tbl <- round(AgeSex_tbl,digits=2)
AgeSex_tbl 
```

<font size="4">

|       Insights : Over 91 percent of students are under age 19, and nearly 60 percent of students are female.

<br/>

|         <font size="5">- Average grade by gender
```{r AveGrade_by_sex}
Ave_grade_sex <- students %>%
  group_by(sex)%>%
  summarise(mean_grade_g1=mean(G1),mean_grade_g2=mean(G2),mean_grade_g3=mean(G3))
Ave_grade_sex

```
<font size="4">

|       Insights : No gender gaps are found in academic grades.


<br/>
---

Since the age composition in this data is unique, I need to be careful with my analysis that includes the 'age' attribute. I would explore for more findings. 

|         <font size="5">
### - Alcohol consumption by gender and age {.tabset}


```{r AgeSex_alcohol_table, warning=FALSE, message=FALSE, warning=FALSE}
library(ggplot2)

# create table and plot that show alcohol consumption by age
# Dalc : workday alcohol consumption, Walc : weekend alcohol consumption (1:very low - 5:very high)

Alcohol_AgeSex <- students %>%
  group_by(age,sex)%>%
  summarise(mean_dalc=mean(Dalc),mean_walc=mean(Walc))
Alcohol_AgeSex

```

#### Workday

```{r}
ggplot(Alcohol_AgeSex, aes(x=age, y=mean_dalc, colour = sex)) +
  labs(title = "Alcohol Consumption (Workday) by Age & Sex",
       x ='Age', y ='Mean Alcohol Consumption')+
  scale_colour_discrete(name = "Gender", labels = c("Female", "Male"))+
  geom_line()
```


#### Weekend
```{r}
ggplot(Alcohol_AgeSex, aes(x=age, y=mean_walc, colour = sex)) + 
  labs(title = "Alcohol Consumption (Weekend) by Age & Gender",
       x ='Age', y ='Mean Alcohol Consumption')+
  scale_colour_discrete(name = "Gender", labels = c("Female", "Male"))+
  geom_line()
```

### {.tabset-pills}

---

|         <font size="5">
### - School grade by gender and age {.tabset}


```{r AgeSex_grade_table, warning=FALSE, message=FALSE, warning=FALSE}
# create table and plot that show school grade by age & gender

grade_AgeSex <- students %>%
  group_by(age,sex)%>%
  summarise(mean_grade1=mean(G1),mean_grade2=mean(G2),mean_grade3=mean(G3))
grade_AgeSex

```

#### 1st period

```{r}
ggplot(grade_AgeSex, aes(x=age, y=mean_grade1, colour = sex)) +
    labs(title = "School Grade by Age & Sex",
       x ='Age', y ='Mean School Grade') +
  scale_colour_discrete(name ='Gender', labels =  c('Female','Male'))+
  geom_line()
```

#### 2nd period

```{r}
ggplot(grade_AgeSex, aes(x=age, y=mean_grade2, colour = sex)) +
    labs(title = "School Grade by Age & Sex",
       x ='Age', y ='Mean School Grade') +
  scale_colour_discrete(name ='Gender', labels =  c('Female','Male'))+
  geom_line()
```

#### 3rd period

```{r}
ggplot(grade_AgeSex, aes(x=age, y=mean_grade3, colour = sex)) +
    labs(title = "School Grade by Age & Sex",
       x ='Age', y ='Mean School Grade') +
  scale_colour_discrete(name ='Gender', labels =  c('Female','Male'))+
  geom_line()
```

### {.tabset-pills}

<font size="4">

---

## Problem

|       The previous charts "Alcohol Consumption by Age & Sex" show that male's alcohol intake increases as age increases, but female doesn't. The charts "School Grade by Age & Sex " show opposite results from the previous charts that male's mean school grades decline with increasing age, which seems to be a correlate movement with alcohol consumption. Since average grades have no gender gaps, these results suggest that age or gender is not likely to be the factor of students' grade, but the amount of drinking would be.
To understand more about students' consumption of alcohol, these questions arise;

* Does alcohol consumption affect students' health, how?
* Does alcohol consumption relate to students' school outcome, how? 

By finding answers for these questions, I am hoping to find a way to help current/future students to control their alcohol amount and achieve their academic goals.  

<br/>

|       For the first question, **does alcohol consumption affect students' health, how?** I would compare students' alcohol amount to their health self-assessment. 

### - Health Status vs Alcohol Consumption {.tabset}
```{r}
#health : health status  (1:very bad - 5:very good)

health_alc <- students %>%
  group_by(health)%>%
  summarise(mean_dalc=mean(Dalc),mean_walc=mean(Walc)) 
health_alc

```


#### Workday

```{r health_alcohol_graph, message=FALSE}
ggplot(health_alc, aes(x=mean_dalc, y=health)) + 
  labs(title = "Alcohol Consumption (Workday) vs Health Status",
        x ='Mean Alcohol Consumption (Workday)', y ='Curent Health Status')+
  geom_line()


```

#### Workday
```{r}

ggplot(health_alc, aes(x=mean_walc, y=health)) + 
  labs(title = "Alcohol Consumption (Weekend) vs health Status",
       x ='Mean Alcohol Consumption (Weekend) Workday', y ='Current Health Status')+
  geom_line()
```

### {.tabset-pills}

<font size="4">

|       Insights : The graphs show a positive relationship between drinking and health status. Being healthy drives higher consumption of alcohol, or alcohol leads healthy body? Generally, the latter would be more possible, but I cannot conclude here because the data doesn't include a long term aspect. So the answer for the first question is **alcohol consumption and students' health seem to relate each other, but this data doesn't include enough facts to bring any meaningful conclusion**.

<br/>

|       The second question, **does alcohol consumption relate to students' school outcome, how?** would be answered by comparing students' grade to their drinking amount.

### - Alcohol consumption vs Grade {.tabset}
```{r grade_alcohol_table, message=FALSE}
AveGrade_Dalc <- students %>%
  group_by(Dalc)%>%
  summarise(mean_grade_g1=mean(G1),mean_grade_g2=mean(G2),mean_grade_g3=mean(G3))
AveGrade_Dalc

AveGrade_Walc <- students %>%
  group_by(Walc)%>%
  summarise(mean_grade_g1=mean(G1),mean_grade_g2=mean(G2),mean_grade_g3=mean(G3))
AveGrade_Walc
```

#### Workday
```{r grade_alcohol_graph, message=FALSE, warning=FALSE}
ggplot(AveGrade_Dalc, aes(x=Dalc)) + 
  geom_line(aes(y = mean_grade_g1, color = "1st")) + 
  geom_line(aes(y = mean_grade_g2, color = "2nd")) +
  geom_line(aes(y = mean_grade_g3, color = "3rd")) +
  labs(title = "Alcohol Consumption (Workday) vs Grade",
       color = 'Grading Period',tag = "A") +
  xlab("Alcohol Consumption (Workday)") + 
  ylab("Mean Grade")+ 
  scale_color_manual(values = c("lightsalmon", "mediumspringgreen","deepskyblue"))


```

#### Weekend
```{r}

ggplot(AveGrade_Walc, aes(x=Walc)) + 
  geom_line(aes(y = mean_grade_g1, color = "1st")) + 
  geom_line(aes(y = mean_grade_g2, color = "2nd")) +
  geom_line(aes(y = mean_grade_g3, color = "3rd")) +
  labs(title = "Alcohol Consumption (Weekend) vs Grade",
       color = 'Grading Period',tag = "B") +
  xlab("Alcohol Consumption (Weekend)") + 
  ylab("Mean Grade")+ 
  scale_color_manual(values = c("lightsalmon", "mediumspringgreen","deepskyblue"))

```

### {.tabset-pills}

<font size="4">

|       Insights : It is clear that students' grade and their alcohol intake strongly relate each other.

<br/>

|       The answer for the second question, **does alcohol consumption relate to students' school outcome, how?** can be explain by; **alcohol affects school grade negatively as the graphs show. The more the alcohol intake, the worse the school grade**. Now move on to figure out "how" alcohol affect students. The data has over 25 attributes, and those can be grouped in 5 categories (**a living location, family, goal-orientation, extra classes, and personal lifestyle**). Let's see how each element connects to alcohol consumption.

<br/>

<style>
.nobullet li {
  list-style-type: none;
p {
  text-indent: 1em;
}
}
</style>

<div class="nobullet">
* <font size="5">1. Living location
</div>


### <font size="4"> - Location {.tabset}
#### table
```{r location_alcohol, message=FALSE}
# address: student's home address type (U:urban, R:rural)

Address_alc <- students %>%
  group_by(address)%>%
  summarise(mean_dalc=mean(Dalc),mean_walc=mean(Walc))
Address_alc

Address_alc_m <- melt(Address_alc[,c('address','mean_dalc','mean_walc')], id.vars = 1)

```

#### plot
```{r}

ggplot(Address_alc_m,aes(x=address, y=value))+
  geom_bar(aes(fill=variable),stat = 'identity', position = "dodge") +
  labs(title = "Alcohol Consumption vs Address")+
  xlab("Home Address Type")+ ylab("Mean Alcohol Consumption") +
  scale_fill_discrete(name="Alcohol on:",
                      labels = c("Workday", "Weekend"))+
    scale_x_discrete(labels = c('R'='Rural','U'='Urban'))
```

### {.tabset-pills}

<br/>

###  <font size="4">-  Traveltime {.tabset}
#### table
```{r traveltime_alcohol}
# traveltime : home to school travel time (1: < 15 min, 2: 15-30 min, 3: 30 min-1hour, 4: >1 hour)

tratime_alc <- students %>%
  group_by(traveltime)%>%
  summarise(mean_dalc=mean(Dalc),mean_walc=mean(Walc))
tratime_alc

tratime_alc_m <- melt(tratime_alc[,c('traveltime','mean_dalc','mean_walc')], id.vars = 1)

```

#### plot
```{r}

ggplot(tratime_alc_m,aes(x=traveltime, y=value))+
  geom_bar(aes(fill=variable),stat = 'identity', position = "dodge") +
  labs(title = "Alcohol Consumption vs Travel Time")+
  xlab("Travel Time from Home to School")+ ylab("Mean Alcohol Consumption") +
  scale_fill_discrete(name="Alcohol on:", labels = c("Workday", "Weekend"))
```

### {.tabset-pills}

<font size="4">

|       Insights : From the aspect of students' living location, alcohol intake doesn't matter either in urban or rural, but it goes up as travel time increases.

<br/>

<div class="nobullet">
* <font size="5">2. Family
</div>

###  <font size="4">- Parents' Education {.tabset}
#### table(father)
```{r Fedu_alcohol}
# Fedu: Father's education (0: none, 1: primary education(4th grade), 2: 5 -9th grade, 3: secondary education, 4:higher education)

fedu_alc <- students %>%
  group_by(Fedu)%>%
  summarise(mean_dalc=mean(Dalc),mean_walc=mean(Walc))
fedu_alc

fedu_alc_m <- melt(fedu_alc[,c('Fedu','mean_dalc','mean_walc')], id.vars = 1)
```

#### plot(father)
```{r}
ggplot(fedu_alc_m,aes(x=Fedu, y=value))+
  geom_bar(aes(fill=variable),stat = 'identity', position = "dodge") +
  labs(title = "Alcohol Consumption vs Travel Time")+
  xlab("Father's Education")+ ylab("Mean Alcohol Consumption") +
  scale_fill_discrete(name="Alcohol on:", labels = c("Workday", "Weekend"))

```

#### table(mother)
```{r}
# Medu : Mother's education

medu_alc <- students %>%
  group_by(Medu)%>%
  summarise(mean_dalc=mean(Dalc),mean_walc=mean(Walc))
medu_alc

medu_alc_m <- melt(medu_alc[,c('Medu','mean_dalc','mean_walc')], id.vars = 1)
```

#### plot(mother)
```{r}
ggplot(medu_alc_m,aes(x=Medu, y=value))+
  geom_bar(aes(fill=variable),stat = 'identity', position = "dodge") +
  labs(title = "Alcohol Consumption vs Travel Time")+
  xlab("Mather's Education")+ ylab("Mean Alcohol Consumption") +
  scale_fill_discrete(name="Alcohol on:", labels = c("Workday", "Weekend"))

```

### {.tabset-pills}

<br/>

### <font size="4">- Parents' job {.tabset}
#### table(father)
```{r}
# Fjob : Father's job ('teacher', 'health' care related, civil 'services' (e.g. administrative or police), 'at_home' or 'other')  

fjob_alc <- students %>%
  group_by(Fjob)%>%
  summarise(mean_dalc=mean(Dalc), mean_walc=mean(Walc))
fjob_alc

fjob_alc_m <- melt(fjob_alc[,c('Fjob','mean_dalc','mean_walc')], id.vars = 1)

```
#### plot(father)
```{r}
ggplot(fjob_alc_m,aes(x=Fjob, y=value))+
  geom_bar(aes(fill=variable),stat = 'identity', position = "dodge") +
  labs(title = "Alcohol Consumption vs Travel Time")+
  xlab("Father's Job")+ ylab("Mean Alcohol Consumption") +
  scale_fill_discrete(name="Alcohol on:", labels = c("Workday", "Weekend"))+
  theme(axis.text.x = element_text(angle=45))
 
```

#### table(mother)
```{r}
# Mother's job (Mjob)
 
mjob_alc <- students %>%
  group_by(Mjob)%>%
  summarise(mean_dalc=mean(Dalc),mean_walc=mean(Walc))
mjob_alc

mjob_alc_m <- melt(mjob_alc[,c('Mjob','mean_dalc','mean_walc')], id.vars = 1)

```
#### plot(father)
```{r}
ggplot(mjob_alc_m,aes(x=Mjob, y=value))+
  geom_bar(aes(fill=variable),stat = 'identity', position = "dodge") +
  labs(title = "Alcohol Consumption vs Travel Time")+
  xlab("Mother's Job")+ ylab("Mean Alcohol Consumption") +
  scale_fill_discrete(name="Alcohol on:", labels = c("Workday", "Weekend"))+
  theme(axis.text.x = element_text(angle=45))
 
```

### {.tabset-pills}

<br/>

### <font size='4'> - Student's guardian {.tabset}
#### table
```{r}
# guardian : Student's guardian  ('mother', 'father' or 'other')

guard_alc <- students %>%
  group_by(guardian)%>%
  summarise(mean_dalc=mean(Dalc),mean_walc=mean(Walc))
guard_alc

guard_alc_m <- melt(guard_alc[,c('guardian','mean_dalc','mean_walc')], id.vars = 1)
```

#### plot
```{r}
ggplot(guard_alc_m,aes(x=guardian, y=value))+
  geom_bar(aes(fill=variable),stat = 'identity', position = "dodge") +
  labs(title = "Alcohol Consumption vs Guradian")+
  xlab("Guardian")+ ylab("Mean Alcohol Consumption") +
  scale_fill_discrete(name="Alcohol on:", labels = c("Workday", "Weekend"))

```

### {.tabset-pills}

<br/>

### <font size='4'> - Cohabitation with parents{.tabset}
#### table
```{r Pstatus_alcohol}
# Pstatus : parent's cohabitation with parent (T: living together, A: apart)

psta_alc <- students %>%
  group_by(Pstatus)%>%
  summarise(mean_dalc=mean(Dalc),mean_walc=mean(Walc))
psta_alc

psta_alc_m <- melt(psta_alc[,c('Pstatus','mean_dalc','mean_walc')], id.vars = 1)

```

#### plot
```{r}

ggplot(psta_alc_m,aes(x=Pstatus, y=value))+
  geom_bar(aes(fill=variable),stat = 'identity', position = "dodge") +
  labs(title = "Alcohol Consumption vs Cohibitation")+
  xlab("Cohabitation with Parents")+ ylab("Mean Alcohol Consumption") +
  scale_fill_discrete(name="Alcohol on:", labels = c("Workday", "Weekend"))+
  scale_x_discrete(labels = c('T'= 'Living Together', 'A' = 'Apart'))

```

### {.tabset-pills}

<br/>

### <font size='4'> - Family size{.tabset}
#### table
```{r}
# Size of family binary: 'LE3' - less or equal to 3 or 'GT3' - greater than 3) (famsize)

fansize_alc <- students %>%
  group_by(famsize)%>%
  summarise(mean_dalc=mean(Dalc),mean_walc=mean(Walc))
fansize_alc

fansize_alc_m <- melt(fansize_alc[,c('famsize','mean_dalc','mean_walc')], id.vars = 1)

```

#### plot
```{r}
ggplot(fansize_alc_m,aes(x=famsize, y=value))+
  geom_bar(aes(fill=variable),stat = 'identity', position = "dodge") +
  labs(title = "Alcohol Consumption vs Family Size")+
  xlab("Family Size")+ ylab("Mean Alcohol Consumption") +
  scale_fill_discrete(name="Alcohol on:", labels = c("Workday", "Weekend"))+
  scale_x_discrete(labels = c('LE3'= 'Less or Equal to 3', 'GT3' ='Greater than 3'))

```

### {.tabset-pills}

<br/>

### <font size='4'> - Relationship with family{.tabset}
#### table
```{r}
# Famrel : Relationship with family (From 1:very bad to 5:excellent) 

famrel_alc <- students %>%
  group_by(famrel)%>%
  summarise(mean_dalc=mean(Dalc),mean_walc=mean(Walc))
famrel_alc

famrel_alc_m <- melt(famrel_alc[,c('famrel','mean_dalc','mean_walc')], id.vars = 1)

```

#### plot
```{r}
ggplot(famrel_alc_m,aes(x=famrel, y=value))+
  geom_bar(aes(fill=variable),stat = 'identity', position = "dodge") +
  labs(title = "Alcohol Consumption vs Family Relationship")+
  xlab("Relationship with Family")+ ylab("Mean Alcohol Consumption") +
  scale_fill_discrete(name="Alcohol on:", labels = c("Workday", "Weekend"))

```

### {.tabset-pills}

<br/>

### <font size='4'> - Educational support from family{.tabset}
#### table
```{r}
# famsup : Educational support from family (yes or no)

famsup_alc <- students %>%
  group_by(famsup)%>%
  summarise(mean_dalc=mean(Dalc),mean_walc=mean(Walc))
famsup_alc

famsup_alc_m <- melt(famsup_alc[,c('famsup','mean_dalc','mean_walc')], id.vars = 1)

```

#### plot
```{r}
ggplot(famsup_alc_m,aes(x=famsup, y=value))+
  geom_bar(aes(fill=variable),stat = 'identity', position = "dodge") +
  labs(title = "Alcohol Consumption vs Educational Support from Family")+
  xlab("Family Support")+ ylab("Mean Alcohol Consumption") +
  scale_fill_discrete(name="Alcohol on:", labels = c("Workday", "Weekend"))

```

### {.tabset-pills}

<font size="4">

|       Insights : Students who live with parents, and have family members equal or less than 3 tend to intake more alcohol than people who are in the opposite situation. On the other hand, students having a strong family relationship and getting family support tend to keep their alcohol intake low.            

<br/>

<div class="nobullet">
* <font size="5">3. Goal-orientation
</div>

### <font size='4'> - Number of school absences  {.tabset}

#### table
```{r}
# number of school absences (from 0 to 93) (absences)

absence_alc_all <- students %>% 
  select(absences,Dalc, Walc)
head(absence_alc_all)

```


#### plot
```{r}

ggplot(absence_alc_all,aes(x=absences))+
  geom_jitter(aes(y = Dalc) , color = 'dodgerblue3') +
  geom_jitter(aes(y = Walc) , color = 'lightpink3') +
  labs(title = "Alcohol Consumption vs Absences")+
  xlab("Absence")+ ylab("Mean Alcohol Consumption") +
  scale_fill_discrete(name="Alcohol on:",
                      labels=c("Workday", "Weekend"))
```

### {.tabset-pills}

<br/>

### <font size='4'> - Number of past class failures {.tabset}
#### table
```{r}
# failures : number of past class failures (n if 1<=n<3, else 4)
fail_alc_all <- students %>% 
  select(failures,Dalc, Walc)
head(fail_alc_all)
```

#### plot
```{r absence_failures_graph}
ggplot(fail_alc_all,aes(x=failures))+
  geom_jitter(aes(y = Dalc) , color = 'dodgerblue3') +
  geom_jitter(aes(y = Walc) , color = 'lightpink3') +
  labs(title = "Alcohol Consumption vs Failures")+
  xlab("Failure")+ ylab("Mean Alcohol Consumption") +
  scale_fill_discrete(name="Alcohol on:",
                      labels=c("Workday", "Weekend"))
```

### {.tabset-pills}

<br/>

### <font size='4'> - Aspiration for higher education {.tabset}
#### table

```{r}
# higher : Aspiration for higher education (yes or no)
high_alc <- students %>%
  group_by(higher)%>%
  summarise(mean_dalc=mean(Dalc),mean_walc=mean(Walc))
high_alc

high_alc_m <- melt(high_alc[,c('higher','mean_dalc','mean_walc')], id.vars = 1)
```

#### plot
```{r}
ggplot(high_alc_m,aes(x=higher, y=value))+
  geom_bar(aes(fill=variable),stat = 'identity', position = "dodge") +
  labs(title = "Alcohol Consumption vs Higher Education")+
  xlab("Higher Education")+ ylab("Mean Alcohol Consumption") +
  scale_fill_discrete(name="Alcohol on:", labels = c("Workday", "Weekend"))
```

### {.tabset-pills}

<br/>

### <font size='4'> - Weekly study time {.tabset}
#### table
```{r}
# studytime : weekly study time (1: <2 hours, 2: 2-5 hours, 3: 5-10 hours, 4: >10 hours) 

stu_time_alc <- students %>%
  group_by(studytime)%>%
  summarise(mean_dalc=mean(Dalc),mean_walc=mean(Walc))
stu_time_alc

stu_time_alc_m <- melt(stu_time_alc[,c('studytime','mean_dalc','mean_walc')], id.vars = 1)

```


#### plot
```{r absence_studytime_graph}
ggplot(stu_time_alc_m,aes(x=studytime, y=value))+
  geom_bar(aes(fill=variable),stat = 'identity', position = "dodge") +
  labs(title = "Alcohol Consumption vs Study Time")+
  xlab("Study Time")+ ylab("Mean Alcohol Consumption") +
  scale_fill_discrete(name="Alcohol on:", labels = c("Workday", "Weekend"))

```

### {.tabset-pills}

<font size="4">

|       Insights : Drinking doesn't seem to relate to the number of school absences and class failures. However, students who want to take higher education and study harder (especially studying for 5 - 10 hours/week) have the least amount of alcohol of all. Note that studying over 10 hours increases the level of drinking.

<br/>

<div class="nobullet">
* <font size="5">4. Extra classes
</div> 

### <font size='4'> - Extra curricular activities {.tabset}
#### table
```{r}
# activities : extra curricular activities (yes or no)

active_curr_alc <- students %>%
  group_by(activities)%>%
  summarise(mean_dalc=mean(Dalc),mean_walc=mean(Walc))
active_curr_alc

active_curr_alc_m <- melt(active_curr_alc[,c('activities','mean_dalc','mean_walc')], id.vars = 1)
```

#### plot
```{r}
ggplot(active_curr_alc_m,aes(x=activities, y=value))+
  geom_bar(aes(fill=variable),stat = 'identity', position = "dodge") +
  labs(title = "Alcohol Consumption vs Extra Curricular Activities")+
  xlab("Extra Curricular Activities")+ ylab("Mean Alcohol Consumption") +
  scale_fill_discrete(name="Alcohol on:", labels = c("Workday", "Weekend"))
```

### {.tabset-pills}

<br/>

### <font size='4'> - Attended nursery school {.tabset}
#### table

```{r}
# nursery : attended nursery school (yes or no)
nursery_alc <- students %>%
  group_by(nursery)%>%
  summarise(mean_dalc=mean(Dalc),mean_walc=mean(Walc))
nursery_alc

nursery_alc_m <- melt(nursery_alc[,c('nursery','mean_dalc','mean_walc')], id.vars = 1)
```

#### plot
```{r}

ggplot(nursery_alc_m,aes(x=nursery, y=value))+
  geom_bar(aes(fill=variable),stat = 'identity', position = "dodge") +
  labs(title = "Alcohol Consumption vs Attending Nursery School")+
  xlab("Attending Nursery School")+ ylab("Mean Alcohol Consumption") +
  scale_fill_discrete(name="Alcohol on:", labels = c("Workday", "Weekend"))
```

### {.tabset-pills}

<br/>

### <font size='4'> - Extra paid classes within the course subject {.tabset}
#### table

```{r}
# Paid : extra paid classes within the course subject (Math or Portuguese) (yes or no)
paid_alc <- students %>%
  group_by(paid)%>%
  summarise(mean_dalc=mean(Dalc),mean_walc=mean(Walc))
paid_alc

paid_alc_m <- melt(paid_alc[,c('paid','mean_dalc','mean_walc')], id.vars = 1)
```

#### plot
```{r}
ggplot(paid_alc_m,aes(x=paid, y=value))+
  geom_bar(aes(fill=variable),stat = 'identity', position = "dodge") +
  labs(title = "Alcohol Consumption vs Paid classes")+
  xlab("Paid classes")+ ylab("Mean Alcohol Consumption") +
  scale_fill_discrete(name="Alcohol on:", labels = c("Workday", "Weekend"))
```

### {.tabset-pills}

<br/>

### <font size='4'> - Extra educational support from school {.tabset}
#### table
```{r}
# scholsup : extra educational support (yes or no)
schoolsup_alc <- students %>%
  group_by(schoolsup)%>%
  summarise(mean_dalc=mean(Dalc),mean_walc=mean(Walc))
schoolsup_alc

schoolsup_alc_m <- melt(schoolsup_alc[,c('schoolsup','mean_dalc','mean_walc')], id.vars = 1)
```

#### plot
```{r}
ggplot(schoolsup_alc_m,aes(x=schoolsup, y=value))+
  geom_bar(aes(fill=variable),stat = 'identity', position = "dodge") +
  labs(title = "Alcohol Consumption vs Extra Educational Support")+
  xlab("Extra Educational Support")+ ylab("Mean Alcohol Consumption") +
  scale_fill_discrete(name="Alcohol on:", labels = c("Workday", "Weekend"))
```

### {.tabset-pills}

<font size="4">

|       Insights : It seems extra classes and curricula can relate to higher alcohol consumption. Like students who study over 10 hours, drinking might relate to students' stress level, while students who get extra educational support from school have lower drinking level than students without the support.

<br/>

<div class="nobullet">
* <font size="5">5. Personal Lifestyle
</div> 


### <font size='4'> - Going out with friends {.tabset}
#### table

```{r}
# goout : going out with friends (1:very low - 5:very high)
goout_alc <- students %>%
  group_by(goout)%>%
  summarise(mean_dalc=mean(Dalc),mean_walc=mean(Walc))
goout_alc

goout_alc_m <- melt(goout_alc[,c('goout','mean_dalc','mean_walc')],id.vars = 1)

```

#### barplot
```{r}
ggplot(goout_alc_m,aes(x=goout, y=value))+
  geom_bar(aes(fill=variable),stat = 'identity', position = "dodge") +
  labs(title = "Alcohol Consumption vs Going Out")+
  xlab("Frequency of Going Out")+ ylab("Mean Alcohol Consumption") +
  scale_fill_discrete(name="Alcohol on:",
                        labels=c("Workday", "Weekend"))

```

### {.tabset-pills}

<br/>

### <font size='4'> - Romantic Relationship {.tabset}
#### table

```{r}
# romantic : with a romantic relationship (yes or no)
romantic_alc <- students %>%
  group_by(romantic)%>%
  summarise(mean_dalc=mean(Dalc),mean_walc=mean(Walc))
romantic_alc

romantic_alc_m <- melt(romantic_alc[,c('romantic','mean_dalc','mean_walc')],id.vars = 1)

```

#### barplot
```{r}
ggplot(romantic_alc_m,aes(x=romantic, y=value))+
  geom_bar(aes(fill=variable),stat = 'identity', position = "dodge") +
  labs(title = "Alcohol Consumption vs Romantic Relationship")+
  xlab("With a Romantic Relationship")+ ylab("Mean Alcohol Consumption") +
  scale_fill_discrete(name="Alcohol on:",
                        labels=c("Workday", "Weekend"))

```

### {.tabset-pills}


<br/>

### <font size='4'> - Free time after school {.tabset}
#### table
```{r}
# freetime : free time after school (1:very low - 5: very high)
free_alc <- students %>%
  group_by(freetime)%>%
  summarise(mean_dalc=mean(Dalc),mean_walc=mean(Walc))
free_alc

free_alc_m <- melt(free_alc[,c('freetime','mean_dalc','mean_walc')],id.vars = 1)

```

#### barplot
```{r}
ggplot(free_alc_m,aes(x=freetime, y=value))+
  geom_bar(aes(fill=variable),stat = 'identity', position = "dodge") +
  labs(title = "Alcohol Consumption vs Free Time")+
  xlab("Free Time")+ ylab("Mean Alcohol Consumption") +
  scale_fill_discrete(name="Alcohol on:",
                        labels=c("Workday", "Weekend"))

```

### {.tabset-pills}

<br/>

Additionally, see how study time relate to free time and school grade.

### <font size='4'> - Free time vs Study Time {.tabset}
#### table
```{r}
# freetime : free time after school (1:very low - 5: very high)
# studytime : weekly study time (1: <2 hours, 2: 2-5 hours, 3: 5-10 hours, 4: >10 hours) 

free_study_mean <- students %>%
    group_by(studytime) %>%
    summarise(mean_freetime=mean(freetime))
free_study_mean

```

#### plot
```{r}
ggplot(free_study_mean, aes(x=studytime)) +
  geom_line(aes(y = mean_freetime), color = 'maroon')+
  labs(title = "Free Time vs Study Time")+
  xlab("Study Time")+ ylab("Mean Free Time")

```

### {.tabset-pills}

<br/>

### <font size='4'> - Study Time vs School Grade {.tabset}
#### table
```{r}
# studytime : weekly study time (1: <2 hours, 2: 2-5 hours, 3: 5-10 hours, 4: >10 hours) 

mean_grade_study <- students %>%
  group_by(studytime)%>%
  summarise(mean_grade1=mean(G1),mean_grade2=mean(G2),mean_grade3=mean(G3))
mean_grade_study

```

#### plot
```{r}
ggplot(mean_grade_study, aes(x = studytime))+
  geom_line(aes(y = mean_grade1, colour = 'pink')) +
  geom_line(aes(y = mean_grade2, colour = 'green')) +
  geom_line(aes(y = mean_grade3, colour = 'blue')) +
  labs(title = "School Grade vs Study Time",
       x ='Study Time (hours)', y ='Mean School Grade') +
  scale_colour_discrete(name ='School Period', labels =  c('1st','2nd','3rd'))+
  scale_x_discrete(limit = c('<2', '2-5','5-10','>10'))
```

### {.tabset-pills}


|       Insights : In personal lifestyle aspects, going out with friends most likely triggers drinking, and having more free time also has a positive relationship with alcohol. Since free time and study time are inversely proportional each other, time management affects students' alcohol consumption and their academic performance.

<br/>

---

## Solution

|       Over all, factors that highly relate to drinking are;

* Having long travel time from home to school (over 30 mins in this case)
* Living with parents and have less family members
* Having poor relationship with one's own family and not getting enough help from them
* Don't have higher educational goals
* Having long study hours (over 10 hours per week in this case)
* Taking extra curricula and classes
* Not getting extra educational support from school
* Having more free time, especially going out with friends often

Knowing all these factors can be useful, but some of them are not practical information for students because some of them (e.g., living location and family related matters) aren't completely their choices. So, the first three bullets are mainly for students parents. At this point, for students, I would recommend to focus on these;

* Having higher goals can lead higher performance
* Manage your study hours and the number of extra classes with one's stress level (This can be challenging, because sometimes students don't have a choice, or they want to despite of their capacity)
* Take advantage of school counselors
* Manage one's free time (the frequency of going out with friends) in a rational way.

---

## Conclusion

|       Alcohol negatively affects students' school performance. Each factor (a living location, family, goal-orientation, extra classes, and personal lifestyle) connects to drinking somehow. Some factors can be handled by students themselves, but others should be supported by their family members. It is important for students and their family members to understand what factors more likely relate to drinking, and by controlling alcohol can improve their academic performance.

---

## Next steps


|       If one aims to be a successful student, they have to at least control their alcohol intake. Things that students want to manage are;

* Watch one's own stress level (over studying like long hours and extra classes have higher possibility of drinking)
* Get support from others, either or both school counselor and family members effectively (for both students' mental and educational aspects)
* Balance study time and free time (fun time helps to release stress, but too much can be counterproductive)

Each student has a different personality, so they have to find their own balance. By keep trying, eventually everyone will know their best balance. 